Diffantom: a whole-brain diffusion MRI phantom derived from a real dataset

### 2015-08-12 paper/main.tex
close #34

\noindent\textbf{\textit{Microstructural model\textcolon}\label{sec:data_model}} %
The simulation process relies on a microstructural model derived from real data.
On one hand, the software requires up to five fraction maps \tmaps{} of
  free- and hindered- diffusion (see \autoref{fig:figure1}, box A).
These compartments will be derived from the macroscopic structure of tissues within the brain,
  specified in the following order~%
\footnote{Corresponding to the \emph{5TT format} established with the latest version 3.0
  of \emph{MRTrix} \citep{tournier_mrtrix_2012}}:
\gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue~%
\footnote{Since here we simulate healthy subjects, the last fraction map $F_5$ is empty and can be omitted}.
On the other hand, the restricted-diffusion compartments are specified by up to three volume fractions \fmaps{}
  of three single fiber populations per voxel along with their corresponding direction maps \vmaps{}.

@@

\noindent\textbf{\textit{Diffusion signal generation\textcolon}\label{sec:data_dwi}} %
Once a microstructural model of the volume has been synthesized, a map of \glspl{fod} is computed using spherical
  convolution of the single fiber response and the fiber orientation maps $\{\mathbf{V}_i\}$, weighted by the
  corresponding fiber-fractions $\{F_i\}$.
A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:figure1}.
The single fiber response is a Gaussian diffusion tensor with axial symmetry and eigenvalues $\lambda_1=$ 2.2\e{-3}
  \diffunits{} and $\lambda_{2,3}=$ 0.2\e{-3} \diffunits{}.
The resulting \glspl{fod} map is then combined with the isotropic compartments corresponding to $\{T_j\}$
  before generating the signal using \emph{Phantomas} \citep{caruyer_phantomas_2014}.
By default, diffusion data are generated using a scheme of 100 directions distributed in one shell with uniform
  coverage \citep{caruyer_design_2013}.
Custom one- or multi-shell schemes can be generated supplying the tables of corresponding vectors and $b$-values.
Rician noise is also included in \emph{Phantomas}, and the \gls*{snr} can be set by the user.
The default value for \gls*{snr} is preset to 30.0.
\Gls*{csf} was simulated with a unique compartment with isotropic diffusivity $D_{CSF}$ of 3.0\e{-3} \diffunits{}.
Conversely, \Gls*{wm} and \gls*{gm} included one isotropic compartment with $D_{WM} =$ 2.0\e{-4}, $D_{cGM} =$ 7.0\e{-4}
  and $D_{dGM} =$ 9.0\e{-4}, respectively [\diffunits{}].
All these values for diffusivity (and the corresponding to the single-fiber response) can be modified by the user with
  custom settings.

>>>>

\noindent\textbf{\textit{Microstructural model\textcolon}\label{sec:data_model}} %
The simulation process relies on a microstructural model derived from real data.
On one hand, the software requires up to five fraction maps \tmaps{} of
  free- and hindered- diffusion (see \autoref{fig:figure1}, box A).
These compartments will be derived from the macroscopic structure of tissues within the brain,
  specified in the following order\footnote{Corresponding to the \emph{5TT format} established
  with the latest version 3.0 of \emph{MRTrix} \citep{tournier_mrtrix_2012}}:
  \gls*{cgm}, \gls*{dgm}, \gls*{wm}, \gls*{csf}, and abnormal tissue\footnote{Since here we
  simulate healthy subjects, the last fraction map $T_5$ is empty and can be omitted}.
On the other hand, the restricted-diffusion compartments are specified by up to three volume fractions \fmaps{}
  of three single fiber populations per voxel along with their corresponding direction maps \vmaps{}.

@@

\noindent\textbf{\textit{Diffusion signal generation\textcolon}\label{sec:data_dwi}} %
Once a microstructural model of the volume has been synthesized, a map of \glspl{fod} is computed using spherical
  convolution of the single fiber response and the fiber orientation maps $\{\mathbf{V}_i\}$, weighted by the
  corresponding fiber-fractions $\{F_i\}$.
A close-up showing how the \glspl{fod} map looks is presented in \autoref{fig:figure1}.
The single fiber response is a Gaussian diffusion tensor with axial symmetry and eigenvalues $\lambda_1=$ 2.2\e{-3}
  \diffunits{} and $\lambda_{2,3}=$ 0.2\e{-3} \diffunits{}.
The resulting \glspl{fod} map is then combined with the free- and hindered-diffusion compartments corresponding to $\{T_j\}$.
The free-diffusion compartment corresponds to the \gls*{csf} fraction map $T_4$ and is modeled with isotropic
  diffusivity $D_{CSF}$ of 3.0\e{-3} \diffunits{}.
The hindered-diffusion compartments correspond to $\{T_1,T_2,T_3\}$ and are also modeled with isotropic diffusivity
  $D_{WM} =$ 2.0\e{-4}, $D_{cGM} =$ 7.0\e{-4} and $D_{dGM} =$ 9.0\e{-4}, respectively [\diffunits{}].
All these values for diffusivity (and the corresponding to the single-fiber response) can be modified by the user with
  custom settings.
The restricted- and hindered- compartments are then fed into \emph{Phantomas} \citep{caruyer_phantomas_2014}
  and the final \gls*{dmri} signal is obtained.
By default, diffusion data are generated using a scheme of 100 directions distributed in one shell with uniform
  coverage \citep{caruyer_design_2013}.
Custom one- or multi-shell schemes can be generated supplying the tables of corresponding vectors and $b$-values.
Rician noise is also included in \emph{Phantomas}, and the \gls*{snr} can be set by the user.
The default value for \gls*{snr} is preset to 30.0.
